---
title: "DATA 607 - Project 1 - Data Analysis"
author: "Preston Peck"
date: "9/19/2021"
output: html_document
---

# Load Libraries

```{r}
library(dplyr)
library(stringr)
library(readr)
library(purrr)
```

# Import Data

```{r}
tournamentInfo <- read_delim("data/tournamentinfo.txt", "\\n", col_names = FALSE, show_col_types = FALSE)
tournamentInfo
```

# Format Lines
We begin to shape our table:

1. Remove row decoration
2. Time leading and trailing whitespace
3. Remove trailing "|" which will create an extra empty bucket when split
4. Split X1 column into sub-columns by "|"

```{r}
divider <- "-----------------------------------------------------------------------------------------"

tournamentInfo <- (tournamentInfo %>%
  filter(!grepl(divider, X1)) %>%
  lapply(function(x) str_trim(x)) %>%
  lapply(function(x) substr(x, 1, nchar(x) - 1)) %>%
  lapply(function(x) str_split(x, "\\|", simplify = TRUE))
)[[1]]

tournamentInfo <- t(apply(tournamentInfo, 1, function(x) str_squish(x)))
tournamentInfo
```

# Extract Column Names
We look at the first 2 rows and extract column names:

1. Get complete column names from single cells
2. Split combined column names and extract
3. Merge 2 part column names across rows

```{r}
columnNames <- data.frame(c1 = tournamentInfo[1,], c2 = tournamentInfo[2,])

idRtgSplit <- str_split(columnNames[2, 2], " / ") %>%
  unlist

twoPartColumnNames <- columnNames[3:10, 1:2]
concatenatedColumns <- paste(twoPartColumnNames$c1, twoPartColumnNames$c2)

columnNames <- c(
  columnNames[1, 1], 
  columnNames[1, 2], 
  columnNames[2, 1], 
  idRtgSplit[1], 
  idRtgSplit[2], 
  unlist(concatenatedColumns)
)

columnNames
```

# Combine Rows

We combine every 2 rows after the header rows since they comprise a single player's record. We split combined values when necessary.

```{r}
mergedTournamentInfo <- matrix(NA, ncol = 13)
mergedTournamentInfo <- as.data.frame(mergedTournamentInfo)
colnames(mergedTournamentInfo) <- columnNames

max <- length(tournamentInfo[,1])

for (x in 3:max - 1) {
  if (x %% 2 == 1) {
    idRtgSplit <- str_split(tournamentInfo[x + 1, 2], " / ") %>%
      unlist
    
    mergedTournamentInfo <- mergedTournamentInfo %>%
      rbind(c(tournamentInfo[x, 1],
          tournamentInfo[x + 1, 1],
          tournamentInfo[x, 2],
          idRtgSplit[1],
          idRtgSplit[2],
          tournamentInfo[x, 3],
          tournamentInfo[x, 4],
          tournamentInfo[x, 5],
          tournamentInfo[x, 6],
          tournamentInfo[x, 7],
          tournamentInfo[x, 8],
          tournamentInfo[x, 9],
          tournamentInfo[x, 10]
        )
      )
  }
}

mergedTournamentInfo <- mergedTournamentInfo %>%
  slice(-c(1))

mergedTournamentInfo
```

# Extract Subset

We capture a subset of information from the previous table to create a new table and extrude it to create an Average Pre Chess Rating of Opponents column. We iterate through a player's opponent ids and jump to the corresponding row for more information about the opponent and split combined value and format data when necessary.

```{r}
columnNames <- c(
  "Player’s Name", 
  "Player’s State", 
  "Total Number of Points", 
  "Player’s Pre-Rating", 
  "Average Pre Chess Rating of Opponents"
)

formattedTournamentInfo <- matrix(NA, ncol = 5)
formattedTournamentInfo <- as.data.frame(formattedTournamentInfo)
colnames(formattedTournamentInfo) <- columnNames

max <- length(mergedTournamentInfo[, 1])

for (x in 1:max) {
  opponentScores <- c()
  
  for (y in 7:13) {
    index <- mergedTournamentInfo[x, y] %>%
      str_replace_all("[^0-9]", "")
    
    opponentPreScore <- as.numeric(((mergedTournamentInfo[index, 5] %>%
      str_split("->") %>%
      unlist)[1] %>%
      str_split("P") %>%
      unlist)[1] %>%
      str_replace_all("[^0-9]", ""))
    
    opponentScores <- opponentScores %>%
      append(opponentPreScore)
  }
  
  meanOpponentScore <- opponentScores[!is.na(opponentScores) & is.numeric(opponentScores)] %>%
    mean %>%
    round(2)
  
  preScore <- as.numeric(((mergedTournamentInfo[x, 5] %>%
      str_split("->") %>%
      unlist)[1] %>%
      str_split("P") %>%
      unlist)[1] %>%
      str_replace_all("[^0-9]", ""))
    
  formattedTournamentInfo <- formattedTournamentInfo %>%
    rbind(c(mergedTournamentInfo[x, 3],
        mergedTournamentInfo[x, 2],
        mergedTournamentInfo[x, 6],
        preScore,
        meanOpponentScore
      )
    )
}

formattedTournamentInfo <- formattedTournamentInfo %>%
  slice(-c(1))

formattedTournamentInfo
```

# Export Data

```{r}
write.csv(formattedTournamentInfo, "data//tournamentinfo.csv", row.names = FALSE)
```
